Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-08-31 manuscript/manuscript.Rmd
updates page numbering for ESA siwtch around

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{todonotes}
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

@@

\begin{centering}
\textbf{\large{Do we need demographic data to forecast population responses to climate change?}}

\textsc{\small{Andrew T. Tredennick\footnote{E-mail: atredenn@gmail.com} and Peter B. Adler}}

\textit{\small{Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322-5230 USA}}

\end{centering}

@@

@@

@@

>>>>

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{todonotes}
   - \usepackage[document]{ragged2e}
   - \pagewiselinenumbers
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

@@

### Title
Do we need demographic data to forecast population responses to climate change?

### Authors 
1. Andrew T. Tredennick (corresponding author: atredenn@gmail.com) 
2. Peter B. Adler (peter.adler@usu.edu)

### Affiliations
1. Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322-5230 USA
2. Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322-5230 USA

\newpage{}

@@

\setcounter{page}{29}

@@

Figure Legends
--------------
Figure 1. Time series of average percent cover over all quadrats for our four focal species: *Bouteloua gracilis* (BOGR), *Hesperostipa comata* (HECO), *Pascopyrum smithii* (PASM), and *Poa secunda* (POSE). Light grey lines show trajectories of individual quadrats. Note the different y-axis scales across panels.

Figure 2. Work flow of the data aggregation, model fitting, and population simulating.

Figure 3. Sensitivity of equilibrium cover simulated from the IPM to each climate scenario applied to individual and combined vital rates. For example, the points associated with G show the median cover from IPM simulations where a climate perturbation is applied only to the growth regression climate covariates. These simulations use mean parameter values for clarity.

Figure 4. Effect of quadrat sample size on the precision (standard deviation) of main climate effect estimates in the QBM. Increasing the number of quadrats results in diminishing returns in terms of parameter certainty. Light dashed lines show individual climate effects at five quadrat sample sizes. Thick dark lines are inverse gaussian fits showing the mean effect of increasing quadrat sample size on parameter precision.

Figure 5. Correlations (r) between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species. Across top panels, 'growth' = growth regression, 'rec' = recruitment regression, 'surv' = survival regression.

Figure 6. Mean (points) and 90% quantiles (errorbars) for the proportional difference between baseline simulations (using observed climate) and the climate pertubation simulation on the x-axis. We calculated proportional difference as log(perturbed climate cover) - log(observed climate cover), where 'perturbed' and 'observed' refer to the climate time series used to drive interannual variation in the simulations. Model error and parameter uncertainty were propagated through the simulation phase. Climate simulations are as in Figure 3.

@@

\setcounter{page}{24}

### 2015-08-31 manuscript/manuscript.tex
updates page numbering for ESA siwtch around

\documentclass[12pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{todonotes}

@@

\normalsize

\begin{centering}
\textbf{\large{Do we need demographic data to forecast population responses to climate change?}}

\textsc{\small{Andrew T. Tredennick\footnote{E-mail: atredenn@gmail.com} and Peter B. Adler}}

\textit{\small{Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322-5230 USA}}

\end{centering}

@@

Our demographic data come from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The
dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata for a complete description. The site is about 800 m above sea
level and mean annual precipitation (1878-2009) is 334 mm, with most
annual precipitation falling from April through September. The community
is grass-dominated and we focused on the four most abundant grass
species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata}
(HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE)
(Fig. 1).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing$\times$climate interactions. The annual maps of the
quadrats were digitized and the fates of individual plants tracked and
extracted using a computer program (Lauenroth and Adler 2008, Chu et al.
2014). Daily climate data are available for the duration of the data
collection period (1932 - 1945) from the Miles City airport, Wiley
Field, 9 km from the study site.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data, we fit regressions for survival, growth, and
recruitment for each species. At the quadrat-level, we fit a single
regression model for population growth. We describe the statistical
models separately since fitting the models required different
approaches. All models contain five climate covariates that we chose
\emph{a priori}: ``water year'' precipitation at \emph{t}-1 (lagppt);
April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and April through June temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions. Climate $\times$ size interactions are for climate
main effects only; we do not include interactions between size and pairs
of interacting climate effects.

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the model likelihood. For the likelihood models, $\textbf{y}^X$ is
always the relevant vector of observations for vital rate \emph{X}
($X = S, G, R, or P$ for survival, growth, recruitment, or population
growth). For example, $\textbf{y}^S$ is a vector of 0's and 1's
indicating whether a genet survives from \emph{t} to \emph{t+1}, or not.

@@

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a
year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope
parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat
group location, and $\theta^{S}_{k}$ is the fixed parameter for the
effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). Note that
the vector of climate covariates (\textbf{C}) includes climate variable
interactions and climate$\times$size interactions. We include
density-dependence by estimating the effect of crowding on the focal
individual by other individuals of the same species. $\omega$ is the
effect of crowding and $w_{t,Q}$ is the crowding experienced by the
focal individual at time $t$ in quadrat group $Q$. We include a
size$\times$crowding interaction effect ($\nu^{S}$).

We modeled growth as a Gaussian process describing genet size at time
$t+1$ as a function of size at $t$ and climate covariates:

@@

where $x$ is log genet size and all other parameters are as described
for the survival regression. We capture non-constant error variance in
growth by modeling the variance around the growth regression
($\varepsilon$) as a nonlinear function of predicted genet size:

@@

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. THerefore, we model recruitment at the
quadrat level: the number of new individuals of species $j$ in quadrat
$q$ recruiting at time $t+1$ as a function of quadrat ``effective
cover'' ($A'$) in the previous year ($t$). Effective cover is a mixture
of observed cover ($A$) in the focal quadrat ($q$) and the mean cover
across the entire group ($\bar{A}$) of $Q$ quadrats in which $q$ is
located:

@@

where $p$ is a mixing fraction between 0 and 1 that is estimated within
the model.

We assume the number of individuals, $y^{R}$, recruiting at time $t+1$
follows a negative binomial distribution:

@@

The statistical approach used to model aggregated data depends on the
type of data collected. We have percent cover data, which can easily be
transformed to proportion data. We first considered fitting three vital
rate models analagous to those we fit at the individual level: one for
probability of extirpation within a quadrat (analagous to survival), one
for cover change within a quadrat (analagous to growth), and one for
probability of colonization within a quadrat (analagous to recruitment).
However, within-quadrat extirpation and colonization events were rare in
our time series ($N=9$ and $N=10$, respectively, across all species).
Given the broad spatial distribution of the quadrats we are studying, it
is safe to assume that these events are in fact rare enough to be
ignored for our purposes. So we constrained our statistical modeling of
vital rates at the population level to change in percent cover within
quadrats. For the remaining discussion of statistical modeling, we refer
to proportion data, which is simply percent cover divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model,
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change from
time $t$ to $t+1$ is

@@

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. 3). Again, note that the climate
covariate vector (\textbf{C}) includes the climate$\times$cover
interaction. The log normal likelihood includes a truncation
(T{[}0,1{]}) to ensure that predicted values do not exceed 100\% cover.

@@

We use a stochastic IPM (Rees and Ellner 2009) that includes the climate
covariates from the vital rate statistical models. In all simulations we
ignore the random year effects so that interannual variation is driven
solely by climate. We fit the random year effects in the vital rate
regressions to avoid over-attributing variation to climate covariates.
Our IPM follows the specification of Chu and Adler (2015) where the
population of species \emph{j} is a density function $n(u_{j},t)$ giving
the density of sized-\emph{u} genets at time \emph{t}. Genet size is on
the natural log scale, so that $n(u_{j},t)du$ is the number of genets
whose area (on the arithmetic scale) is between $e^{u_{j}}$ and
$e^{u_{j}+du}$. The density function for any size \emph{v} at time $t+1$
is

@@

where $k_{j}(v_{j},u_{j},\bar{w_{j}})$ is the population kernel that
describes all possible transitions from size $u$ to $v$ and
$\bar{w_{j}}$ is a scalar representing the average intraspecific
crowding experienced by a genet of size $u_j$ and species $j$. The
integral is evaluated over all possible sizes between predefined lower
(\emph{L}) and upper (\emph{U}) size limits that extend beyond the range
of observed genet sizes.

Since the IPM is spatially-implicit, we cannot calculate neighborhood
crowding for specific genets ($w_{ij}$). Instead, we use an
approximation ($\bar{w_{j}}$) that captures the essential features of
neighborhood interactions (Adler et al. 2010). This approximation relies
on a `no-overlap' rule for conspecific genets to approximate the
overdispersion of large genets in space (Adler et al. 2010).

@@

which means we are calculating growth (\emph{G}) for individuals that
survive (\emph{S}) from time \emph{t} to \emph{t+1} and adding in newly
recruited (\emph{R}) individuals of an average sized one-year-old genet
for the focal species. Our stastical model for recruitment (\emph{R},
described above) returns the number of new recruit produced per quadrat.
Following previous work (Adler et al. 2012, Chu and Adler 2015), we
assume that fecundity increases linearly with size
($R_j(v_{j},u_{j},\bar{w_{j}}) = e^{u_j}R_j(v_{j},\bar{w_{j}})$) to
incorporate the recruitment function in the spatially-implicit IPM.

@@

To simulate our quad-based model (QBM), we simply iterate the
quadrat-level statistical model (Eqs. 9-10). We use the same approach
for drawing parameter values as described for the IPM. After drawing the
appropriate parameter set, we calculate the mean response (population
cover at \emph{t+1} = $x_{t+1}$) according to Eq. 9. We then make a
random draw from a {[}0,1{]} truncated lognormal distribution with mean
equal to $x_{t+1}$ from Eq. 9 and the variance estimate from the fitted
model. We can then project the model forward by drawing a new parameter
set (unique to climate year and MCMC iteration) at each timestep. As
with the IPM, random year effects are ignored for all simulations.

@@

To deterime if the QBM climate effects are correlated with climate
effects for each vital rate model in the IPM, we simply regressed the
QBM climate coefficients against each vital rate model's climate
coefficients and calculated Pearson's $\rho$. Strong correlations
indicate the QBM is capable of detecting climate effects associated with
individual vital rates. A weak correlation indicates the QBM ``misses''
the climate effect on a particular vital rate.

@@

The IPM had lower overall error (MAE, mean absolute error) compared to
the QBM for three species (Table 1). The IPM MAE is significantly lower
at $\alpha=0.05$ for \emph{B. gracilis} (\emph{P} = 0.0012), \emph{H.
comata} (\emph{P} = 4.0586 × 10-8), and \emph{P. smithii} (\emph{P} =
3.183 × 10-5). MAEs are statisticially similar between models for
\emph{P. secunda} (\emph{P} = 0.0922). \emph{P} values are highly
sensitive to sample size, so not entirely appropriate in simulation
exercises where we control the samples size. But, for our purposes they
serve as relatively unbiased comparison metrics. In no case did the QBM
significantly outperform the IPM (Table 1). The IPM was consistently
more precise than the QBM, with lower distances between the 90\%
quantiles across all species (Table 1). In general the IPM outperformed
the QBM because it had (1) lower MAE for three of the four species, (2)
statistically similar MAE for the one other species, and (3)
considerably more precise forecasts for all species.

@@

@@

\subsection{References}\label{references}

@@

Hobbs, N. T., and M. B. Hooten. 2015. Bayesian Models: A Statistical
Primer for Ecologists. Princeton University PressPrinceton.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova,
O'HaraR. B., F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schröder, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}
\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{todonotes}
\usepackage[document]{ragged2e}
\pagewiselinenumbers

@@

\maketitle

\subsubsection{Title}\label{title}

Do we need demographic data to forecast population responses to climate
change?

\subsubsection{Authors}\label{authors}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Andrew T. Tredennick (corresponding author:
  \href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}})
\item
  Peter B. Adler
  (\href{mailto:peter.adler@usu.edu}{\nolinkurl{peter.adler@usu.edu}})
\end{enumerate}

\subsubsection{Affiliations}\label{affiliations}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Department of Wildland Resources and the Ecology Center, 5230 Old Main
  Hill, Utah State University, Logan, Utah 84322-5230 USA
\item
  Department of Wildland Resources and the Ecology Center, 5230 Old Main
  Hill, Utah State University, Logan, Utah 84322-5230 USA
\end{enumerate}

\newpage{}

@@

Our demographic data come from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA (\(46^{\circ}\) 19' N, \(105^{\circ}\) 48' W).
The dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata for a complete description. The site is about 800 m above sea
level and mean annual precipitation (1878-2009) is 334 mm, with most
annual precipitation falling from April through September. The community
is grass-dominated and we focused on the four most abundant grass
species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata}
(HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE)
(Fig. 1).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-\(\text{m}^2\) quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing\(\times\)climate interactions. The annual maps of
the quadrats were digitized and the fates of individual plants tracked
and extracted using a computer program (Lauenroth and Adler 2008, Chu et
al. 2014). Daily climate data are available for the duration of the data
collection period (1932 - 1945) from the Miles City airport, Wiley
Field, 9 km from the study site.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data, we fit regressions for survival, growth, and
recruitment for each species. At the quadrat-level, we fit a single
regression model for population growth. We describe the statistical
models separately since fitting the models required different
approaches. All models contain five climate covariates that we chose
\emph{a priori}: ``water year'' precipitation at \emph{t}-1 (lagppt);
April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and April through June temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 \(\times\) TmeansSpr1) and climate
\(\times\) size interactions. Climate \(\times\) size interactions are
for climate main effects only; we do not include interactions between
size and pairs of interacting climate effects.

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the model likelihood. For the likelihood models, \(\textbf{y}^X\) is
always the relevant vector of observations for vital rate \emph{X}
(\(X = S, G, R, or P\) for survival, growth, recruitment, or population
growth). For example, \(\textbf{y}^S\) is a vector of 0's and 1's
indicating whether a genet survives from \emph{t} to \emph{t+1}, or not.

@@

where \(x_{ij,t}\) is the log of genet size, \(\gamma^{S}_{j,t}\) is a
year-specific intercept, \(\beta^{S}_{j,t}\) is the year-specific slope
parameter for size, \(\phi^{S}_{jQ}\) is the random effect of quadrat
group location, and \(\theta^{S}_{k}\) is the fixed parameter for the
effect of the \(k\)th climate covariate at time \(t\) (\(C_{k,t}\)).
Note that the vector of climate covariates (\textbf{C}) includes climate
variable interactions and climate\(\times\)size interactions. We include
density-dependence by estimating the effect of crowding on the focal
individual by other individuals of the same species. \(\omega\) is the
effect of crowding and \(w_{t,Q}\) is the crowding experienced by the
focal individual at time \(t\) in quadrat group \(Q\). We include a
size\(\times\)crowding interaction effect (\(\nu^{S}\)).

We modeled growth as a Gaussian process describing genet size at time
\(t+1\) as a function of size at \(t\) and climate covariates:

@@

where \(x\) is log genet size and all other parameters are as described
for the survival regression. We capture non-constant error variance in
growth by modeling the variance around the growth regression
(\(\varepsilon\)) as a nonlinear function of predicted genet size:

@@

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. THerefore, we model recruitment at the
quadrat level: the number of new individuals of species \(j\) in quadrat
\(q\) recruiting at time \(t+1\) as a function of quadrat ``effective
cover'' (\(A'\)) in the previous year (\(t\)). Effective cover is a
mixture of observed cover (\(A\)) in the focal quadrat (\(q\)) and the
mean cover across the entire group (\(\bar{A}\)) of \(Q\) quadrats in
which \(q\) is located:

@@

where \(p\) is a mixing fraction between 0 and 1 that is estimated
within the model.

We assume the number of individuals, \(y^{R}\), recruiting at time
\(t+1\) follows a negative binomial distribution:

@@

The statistical approach used to model aggregated data depends on the
type of data collected. We have percent cover data, which can easily be
transformed to proportion data. We first considered fitting three vital
rate models analagous to those we fit at the individual level: one for
probability of extirpation within a quadrat (analagous to survival), one
for cover change within a quadrat (analagous to growth), and one for
probability of colonization within a quadrat (analagous to recruitment).
However, within-quadrat extirpation and colonization events were rare in
our time series (\(N=9\) and \(N=10\), respectively, across all
species). Given the broad spatial distribution of the quadrats we are
studying, it is safe to assume that these events are in fact rare enough
to be ignored for our purposes. So we constrained our statistical
modeling of vital rates at the population level to change in percent
cover within quadrats. For the remaining discussion of statistical
modeling, we refer to proportion data, which is simply percent cover
divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model,
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change from
time \(t\) to \(t+1\) is

@@

where \(x_{jq,t}\) is the log of species' \(j\) proportional cover in
quadrat \(q\) at time \(t\) and all other parameters are as in the
individual-level growth model (Eq. 3). Again, note that the climate
covariate vector (\textbf{C}) includes the climate\(\times\)cover
interaction. The log normal likelihood includes a truncation
(T{[}0,1{]}) to ensure that predicted values do not exceed 100\% cover.

@@

We use a stochastic IPM (Rees and Ellner 2009) that includes the climate
covariates from the vital rate statistical models. In all simulations we
ignore the random year effects so that interannual variation is driven
solely by climate. We fit the random year effects in the vital rate
regressions to avoid over-attributing variation to climate covariates.
Our IPM follows the specification of Chu and Adler (2015) where the
population of species \emph{j} is a density function \(n(u_{j},t)\)
giving the density of sized-\emph{u} genets at time \emph{t}. Genet size
is on the natural log scale, so that \(n(u_{j},t)du\) is the number of
genets whose area (on the arithmetic scale) is between \(e^{u_{j}}\) and
\(e^{u_{j}+du}\). The density function for any size \emph{v} at time
\(t+1\) is

@@

where \(k_{j}(v_{j},u_{j},\bar{w_{j}})\) is the population kernel that
describes all possible transitions from size \(u\) to \(v\) and
\(\bar{w_{j}}\) is a scalar representing the average intraspecific
crowding experienced by a genet of size \(u_j\) and species \(j\). The
integral is evaluated over all possible sizes between predefined lower
(\emph{L}) and upper (\emph{U}) size limits that extend beyond the range
of observed genet sizes.

Since the IPM is spatially-implicit, we cannot calculate neighborhood
crowding for specific genets (\(w_{ij}\)). Instead, we use an
approximation (\(\bar{w_{j}}\)) that captures the essential features of
neighborhood interactions (Adler et al. 2010). This approximation relies
on a `no-overlap' rule for conspecific genets to approximate the
overdispersion of large genets in space (Adler et al. 2010).

@@

which means we are calculating growth (\emph{G}) for individuals that
survive (\emph{S}) from time \emph{t} to \emph{t+1} and adding in newly
recruited (\emph{R}) individuals of an average sized one-year-old genet
for the focal species. Our stastical model for recruitment (\emph{R},
described above) returns the number of new recruit produced per quadrat.
Following previous work (Adler et al. 2012, Chu and Adler 2015), we
assume that fecundity increases linearly with size
(\(R_j(v_{j},u_{j},\bar{w_{j}}) = e^{u_j}R_j(v_{j},\bar{w_{j}})\)) to
incorporate the recruitment function in the spatially-implicit IPM.

@@

To simulate our quad-based model (QBM), we simply iterate the
quadrat-level statistical model (Eqs. 9-10). We use the same approach
for drawing parameter values as described for the IPM. After drawing the
appropriate parameter set, we calculate the mean response (population
cover at \emph{t+1} = \(x_{t+1}\)) according to Eq. 9. We then make a
random draw from a {[}0,1{]} truncated lognormal distribution with mean
equal to \(x_{t+1}\) from Eq. 9 and the variance estimate from the
fitted model. We can then project the model forward by drawing a new
parameter set (unique to climate year and MCMC iteration) at each
timestep. As with the IPM, random year effects are ignored for all
simulations.

@@

To deterime if the QBM climate effects are correlated with climate
effects for each vital rate model in the IPM, we simply regressed the
QBM climate coefficients against each vital rate model's climate
coefficients and calculated Pearson's \(\rho\). Strong correlations
indicate the QBM is capable of detecting climate effects associated with
individual vital rates. A weak correlation indicates the QBM ``misses''
the climate effect on a particular vital rate.

@@

The IPM had lower overall error (MAE, mean absolute error) compared to
the QBM for three species (Table 1). The IPM MAE is significantly lower
at \(\alpha=0.05\) for \emph{B. gracilis} (\emph{P} = 0.0012), \emph{H.
comata} (\emph{P} = 4.0586 × 10-8), and \emph{P. smithii} (\emph{P} =
3.183 × 10-5). MAEs are statisticially similar between models for
\emph{P. secunda} (\emph{P} = 0.0922). \emph{P} values are highly
sensitive to sample size, so not entirely appropriate in simulation
exercises where we control the samples size. But, for our purposes they
serve as relatively unbiased comparison metrics. In no case did the QBM
significantly outperform the IPM (Table 1). The IPM was consistently
more precise than the QBM, with lower distances between the 90\%
quantiles across all species (Table 1). In general the IPM outperformed
the QBM because it had (1) lower MAE for three of the four species, (2)
statistically similar MAE for the one other species, and (3)
considerably more precise forecasts for all species.

@@

\subsection{Figure Legends}\label{figure-legends}

Figure 1. Time series of average percent cover over all quadrats for our
four focal species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa
comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda}
(POSE). Light grey lines show trajectories of individual quadrats. Note
the different y-axis scales across panels.

Figure 2. Work flow of the data aggregation, model fitting, and
population simulating.

Figure 3. Sensitivity of equilibrium cover simulated from the IPM to
each climate scenario applied to individual and combined vital rates.
For example, the points associated with G show the median cover from IPM
simulations where a climate perturbation is applied only to the growth
regression climate covariates. These simulations use mean parameter
values for clarity.

Figure 4. Effect of quadrat sample size on the precision (standard
deviation) of main climate effect estimates in the QBM. Increasing the
number of quadrats results in diminishing returns in terms of parameter
certainty. Light dashed lines show individual climate effects at five
quadrat sample sizes. Thick dark lines are inverse gaussian fits showing
the mean effect of increasing quadrat sample size on parameter
precision.

Figure 5. Correlations (r) between QBM and IPM estimates of climate
effects. We ignore sizeXclimate interactions since these are not
directly comparable across model types. The QBM does not have multiple
vital rates, so its values are repeated across panels within each
species. Across top panels, `growth' = growth regression, `rec' =
recruitment regression, `surv' = survival regression.

Figure 6. Mean (points) and 90\% quantiles (errorbars) for the
proportional difference between baseline simulations (using observed
climate) and the climate pertubation simulation on the x-axis. We
calculated proportional difference as log(perturbed climate cover) -
log(observed climate cover), where `perturbed' and `observed' refer to
the climate time series used to drive interannual variation in the
simulations. Model error and parameter uncertainty were propagated
through the simulation phase. Climate simulations are as in Figure 3.

@@

\setcounter{page}{24}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

@@

Hobbs, N. T., and M. B. Hooten. 2015. Bayesian Models: A Statistical
Primer for Ecologists. Princeton University Press, Princeton.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova, R. B.
O'Hara, F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schröder, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.
